---
title: "Process Historical Temp Data"
author: "Charlotte Mann GSI"
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
mypacks <- c("dplyr", "lubridate", "stringr", "foreach")  # what packages are needed?
lapply(mypacks, library, character.only=TRUE)  # load all packages
```

## Set up
Load scraping functions
```{r, child = '00-wunderground-scrape.Rmd', include = F}
```

Get vector of station codes
```{r}
station_codes <- read.csv("../temp/station_xwalk.csv") %>%
  select(id) %>%
  unlist()
```

Get today's date
```{r}
tdy_date <- Sys.Date()
tdy_date_clean <- str_replace_all(as.character(tdy_date),"\\-","")
```

Pull yesterday's temperature data

```{r}
yst_date <- tdy_date - 1
yst_date_clean <- str_replace_all(as.character(yst_date),"\\-","")

current.temp <- foreach(i = 1:length(station_codes), .combine = rbind) %do% {
  temp <- fetch_wunderground(stationid = station_codes[i], start_date = yst_date_clean)
  summarize_wunderground(temp)
}
```

## Predict from yesterday's temperature for all given future times

```{r}

predict_temp_naive <- function(current.temps, days_lag){
  
  preds <- current.temps %>%
    select(station_code, pred.tmin = min.temp,
           pred.tmax = max.temp, pred.tavg = avg.temp) %>%
    mutate(days_lag = days_lag)
  
  return(preds)
  
}

```

## Predict from historical data - 2 weeks around today's date (local=T) or all days (local = F)


Function to fit a model either on all historical data, or for historical data 2 weeks surrounding date of interest for prediction.

Models are of the form:

$$Y_{s,t,d} = \alpha + \beta Y_{s,t,(d-\ell)} + \alpha_{s}I_s + \beta_{s}Y_{s,t,(d-\ell)}I_s + \epsilon_{s,t,d}$$
Where $s$ is the station id, $t$ is the year, $d$ is the day of the year, and $\ell$ is the number of lag days. $I_s$ is an indicator for the station. 

```{r}
fit_hist_model <- function(hist.temps, days_lag, 
                           local = T, date=NULL){
  
  if(local == T){
    
    start_date = date - weeks(1) - days(days_lag)
      
    startm<-month(start_date)
    startd<-day(start_date)
  
    mod.dat <- hist.temps %>%
      mutate(year = year(date)) %>%
      mutate(start_date = make_date(year, startm, startd)) %>%
      mutate(keep = date %within% interval(start_date, start_date + weeks(2) + days(days_lag)))%>%
      filter(keep) %>%
      group_by(station_code, year) %>%
      mutate(across(tmin:tavg_fill, ~lag(.x, n = days_lag), .names = "{.col}_lag"))
    
  }else{
    
    mod.dat <- hist.temps %>%
      mutate(year = year(date)) %>%
      group_by(station_code, year) %>%
      mutate(across(tmin:tavg_fill, ~lag(.x, n = days_lag), .names = "{.col}_lag"))
    
  } 
    
  tmin.mod <- lm(tmin ~ tmin_lag*station_code, data= mod.dat)
  tmax.mod <- lm(tmax ~ tmax_lag*station_code, data= mod.dat)
  tavg.mod <- lm(tavg_fill ~ tavg_fill_lag*station_code, data= mod.dat)

  return(list(tmin = tmin.mod, tmax = tmax.mod, tavg = tavg.mod))

}
```

Make predictions based on historical data. When predicting only using the weeks surrounding the date,
fit that model. When using all historical data, give the function those model fits (they don't need to be run over and over)

```{r}
predict_temp_hist <- function(current.temps, 
                              local = TRUE,
                              date = NULL, hist.temps = NULL,
                              tmin.model = NULL, tmax.model=NULL, tavg.model=NULL, 
                              days_lag){
  
  if(local){
    mods <- fit_hist_model(date=date, hist.temps=hist.temps, days_lag = days_lag, local=T)
    tmin.model <- mods[["tmin"]]
    tmax.model <- mods[["tmax"]]
    tavg.model <- mods[["tavg"]]
  }
  
  current.temps <- current.temps %>%
    rename(tmin_lag = min.temp,
           tmax_lag = max.temp,
           tavg_fill_lag = mean.temp)
  
  pred.tmin <- predict(tmin.model, newdata = current.temps)
  pred.tmax <- predict(tmax.model, newdata = current.temps)
  pred.tavg <- predict(tavg.model, newdata = current.temps)
  
  preds <- data.frame(station_code = current.temps$station_code,
                      pred.tmin,
                      pred.tmax,
                      pred.tavg,
                      days_lag = days_lag)
  
  return(preds)
} 

```

