---
title: "Process Historical Temp Data"
author: "Charlotte Mann GSI"
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
mypacks <- c("dplyr", "lubridate")  # what packages are needed?
lapply(mypacks, library, character.only=TRUE)  # load all packages
```

Read in raw data, downloaded from NOAA https://www.ncei.noaa.gov/cdo-web/search?datasetid=GHCND. I had to download the data in two different files, which just have different stations.

Data goes from 1/1/1967 through 11/3/2022 (55 years)

```{r}
raw.dat1 <- read.csv("../raw-data/3139726.csv")
raw.dat2 <- read.csv("../raw-data/3139742.csv")
raw.dat <- rbind(raw.dat1, raw.dat2) %>%
  filter(STATION != "USW00023062")

rm(raw.dat1,raw.dat2)
```

Read in manually created station ID xwalk, which matches the NOAA station identifiers to the station codes in wunderground.

```{r}
station.id.xwalk <- read.csv("../temp/station_xwalk.csv")
```

Check that cross walk matches up
```{r}
raw.dat %>%
  select(STATION, NAME,LATITUDE) %>%
  distinct() %>%
  left_join(station.id.xwalk, by = c("STATION", "NAME")) 
```

Clean up historical data column names and generate a lubridate date column

```{r}
hist.dat <- raw.dat %>%
  left_join(station.id.xwalk, by = c("STATION", "NAME")) %>%
  mutate(date = ymd(DATE)) %>%
  select(station_id = STATION,
         station_code = id,
         station_name = NAME,
         station_city = city,
         date_chr = DATE,
         date,
         tavg = TAVG,
         tmin = TMIN,
         tmax = TMAX)
```

```{r}
mean(is.na(hist.dat$tavg))*100
mean(is.na(hist.dat$tmin))*100
mean(is.na(hist.dat$tmax))*100
```

There is a lot of data missing for the average temperature. Fill in with the average of the min and max temperatures in these cases.

```{r}
hist.dat <- hist.dat %>%
  mutate(tavg_fill = ifelse(is.na(tavg), (tmax + tmin)/2, tavg))

mean(is.na(hist.dat$tavg_fill))*100
```

Save processed data

```{r}
save(hist.dat, file = "../temp/hist-dat-clean.Rdata")
```
